---
title: "models_gl"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(patchwork)
library(stargazer)
library(sandwich)
library(Hmisc)
setwd('.')

minimal_theme_GL <- theme(
      axis.text = element_text(color="#959292"),
      axis.line = element_line(color = "#959292", size = .25), 
      axis.title = element_text(color="#959292"),
      axis.ticks = element_line(color = "#959292", size = .25),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.background = element_blank(), 
      plot.title = element_text(color="#959292", size = 11),
      plot.subtitle = element_text(color="#959292"),
      legend.text = element_text(color="#959292"), 
      legend.title = element_blank(), 
      # legend.justification=c(0,1), 
      # legend.position=c(0,1), 
      legend.direction = 'vertical')
```

```{r import data}
data <- read.csv('../data/processed/processed_data.csv')
```

Ensure all of the data still looks as expected. 
```{r}
summary(data)
```

One Issue with NAs

* 363 missing values for `avg_parks_change`

  + Explanation: This occured in cases where a date had NO data for parks changes (at county level), but did have data for retail changes. 

  + Action: will have to drop these observations - don't have way to replace data



## Limited Model
### Only the key variable
```{r model one}
model_one <- data %>%
  lm(avg_retail_rec_change ~ at_home_order, .)
```


## Model Two
### key explanatory variables and covariates that advance modeling goals
```{r model two}
model_two <- data %>%
  lm(avg_retail_rec_change ~ 
       at_home_order +                          # Primary variable of interest
       quarantine_length + mask_order +         # Other covid policies/happenings that may confound
       population_density + new_cases_per_100k, # State level effects that may confound
     .) 
```

## Model Three
### includes covariates from model two, plus a few more - erring on the side of inclusion
```{r model three}
model_three <- data %>%
  lm(avg_retail_rec_change ~ 
       at_home_order +                           # Primary variable of interest
       quarantine_length + mask_order +          # Other covid policies/happenings that may confound
       population_density + new_cases_per_100k + # State level effects that may confound
       new_cases + population +                  # Likely collinear variables
       avg_parks_change,                         # mobility related control we decided against
     .)
```

## Assemble robust standard errors for tests of betas
```{r robust se}
# Variance-covariance matrix for each model
beta_cov1 <- vcovHC.default(model_one, type = "HC3")
beta_cov2 <- vcovHC.default(model_two, type = "HC3")
beta_cov3 <- vcovHC.default(model_three, type = "HC3")

# Extract robust se from vcov matrix diagonal
robust_se1 <- sqrt(diag(beta_cov1))
robust_se2 <- sqrt(diag(beta_cov2))
robust_se3 <- sqrt(diag(beta_cov3))
```


## Summarize models
```{r, results = 'asis', message = FALSE, warning = FALSE, header = FALSE}
stargazer(
  model_one, 
  model_two,
  model_three,
  se = list(robust_se1, robust_se2, robust_se3), # replace classical w/ robust SE
  # type = 'latex'
  type = 'text'
)
```


```{r, normall_distributed_errors}
# describe(data)
data <- data %>% mutate(res = residuals(model_two))
```

```{r, normall_distributed_errors}
hist(x=data$res, xlab='Residuals', main='Histogram of residuals')
qqnorm(y=data$res, pch = 1,frame=FALSE)
qqline(data$res, col = "steelblue", lwd = 2)
```