---
title: "models_gl"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(patchwork)
library(stargazer)
library(sandwich)
setwd('.')

minimal_theme_GL <- theme(
      axis.text = element_text(color="#959292"),
      axis.line = element_line(color = "#959292", size = .25), 
      axis.title = element_text(color="#959292"),
      axis.ticks = element_line(color = "#959292", size = .25),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.background = element_blank(), 
      plot.title = element_text(color="#959292", size = 11),
      plot.subtitle = element_text(color="#959292"),
      legend.text = element_text(color="#959292"), 
      legend.title = element_blank(), 
      # legend.justification=c(0,1), 
      # legend.position=c(0,1), 
      legend.direction = 'vertical')
```

```{r import data}
data_filtered <- read.csv('../data/processed/processed_data.csv') %>%
  mutate(
    start = case_when(
    as.Date(date) >= '2020-03-28'~ 1,
    TRUE ~ 0)
    ) %>%
  filter(as.Date(date) <= '2020-06-15')

data <- read.csv('../data/processed/processed_data.csv') %>%
  mutate(
    start = case_when(
    as.Date(date) >= '2020-03-28'~ 1,
    TRUE ~ 0)
    )

head(data)

```

Next, want to ensure all of the data still looks as expected. 
```{r}
summary(data)
```

Can see that there are two issues in terms of NAs. 

* 363 missing values for `avg_parks_change`

  + Explanation: This occured in cases where a date had NO data for parks changes (at county level), but did have data for retail changes. 

  + Action: will have to drop these observations - don't have way to replace data

<!-- * 910 Missing values for `new_cases` ---- FIXED this in processing file V2 -->

<!--   + Explanation: These are days where no new COVID cases had yet been reported. There was technically no "before" to compare against, so they received NA values. in reality - we can interpret this as being 0 new cases in a day.  -->

<!--   + Action: overwrite All NAs to 0 -->

<!-- Can move on after taking corrective actions -->

## Limited Model
### Only the key variable
```{r}

model_one <- data %>%
  lm(avg_retail_rec_change ~ start*group, .)
summary(model_one)

beta_cov1 <- vcovHC.default(model_one, type = "HC3")
```


## Model Two
### key explanatory variables and covariates that advance modeling goals
```{r}
names(data)
DID <- data_filtered %>%
  lm(avg_retail_rec_change ~ start*group + quarantine_length + mask_order + population_density + new_cases_per_100k, .)

DID2 <- data_filtered %>%
  lm(avg_retail_rec_change ~ start*group + quarantine_length + mask_order + new_cases_per_100k, .)

non_did <- data %>%
  lm(avg_retail_rec_change ~ at_home_order + quarantine_length + mask_order + population_density + new_cases_per_100k, .)


non_did2 <- data %>%
  lm(avg_retail_rec_change ~ at_home_order + quarantine_length + mask_order + new_cases_per_100k, .)

beta_cov0 <- vcovHC.default(DID, type = "HC3")
beta_cov1 <- vcovHC.default(DID2, type = "HC3")
beta_cov2 <- vcovHC.default(non_did, type = "HC3")
beta_cov3 <- vcovHC.default(non_did2, type = "HC3")
```

## Model Three
### includes covariates from model two, plus a few more - erring on the side of inclusion
```{r}
model_three <- data %>%
  lm(avg_retail_rec_change ~ in_quarantine + quarantine_length + mask_order + new_cases +
       avg_parks_change, #+
     # state + date,
    .) # Fixed Effects

beta_cov3 <- vcovHC.default(model_three, type = "HC3")
```

## Summarize models
```{r, results = 'asis', message = FALSE, warning = FALSE, header = FALSE}
# Extract robust se from vcov matrix diagonal
robust_se1 <- sqrt(diag(beta_cov1))
robust_se2 <- sqrt(diag(beta_cov2))
robust_se3 <- sqrt(diag(beta_cov3))

stargazer(
  DID2, 
  DID,
  non_did2,
  non_did,
  # se = list(robust_se0, robust_se1, robust_se2, robust_se3), # load errors from vcovHC
  # type = 'latex'
  type = 'text'
)
```
