---
title: "cleaning"
output: html_document
---

```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(rlang)
```

```{r set graphing theme}
minimal_theme_GL <- theme(
      axis.text = element_text(color="#959292"),
      axis.line = element_line(color = "#959292", size = .25), 
      axis.title = element_text(color="#959292"),
      axis.ticks = element_line(color = "#959292", size = .25),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.background = element_blank(), 
      plot.title = element_text(color="#959292", size = 11),
      plot.subtitle = element_text(color="#959292"),
      legend.text = element_text(color="#959292"), 
      legend.title = element_blank(), 
      # legend.justification=c(0,1), 
      # legend.position=c(0,1), 
      legend.direction = 'vertical')
```


Save Raw Data in DF
```{r pull raw data}
mobility_raw <- read.csv("~/mids_w203_sp_21/lab_2-rbgs/data/raw/2020_US_Region_Mobility_Report.csv")
```

Creating some Funcitons to Aide in the cleaning process
```{r define functions}

count_blank_na <- function(in_df, column_name = in_df[1]) {
  # Purpose: quick count of blanks or NA values in a column
  # Input: df, column_name to be checked 
  # - defaults to first column if none provided
  # Output: number of blanks in column
  
  in_df %>%
  filter({{ column_name }} == "" | 
           is.na({{ column_name }})) %>%
  nrow() %>% cat("\n")
}

add_date_values <- function(in_df, date_field = date) {
  # Purpose: augmenting "date" with common fields - month, year, day of week
  # Input: in_df - dataframe, date_field - name of column with date
  # - date_field will look for "date" column as default
  # Output: in_df, with columns added for month, year, and dow
  
  in_df %>%
    mutate(month = format(as.Date({{date_field}}), "%m"),
         # month = format(as.Date({{date_field}}), "%m-%Y"),
         year = format(as.Date({{date_field}}), "%Y"),
         dow = weekdays(as.Date({{date_field}})))
}

plot_missing_histogram <- function(in_df, var, lod, lab_x = "", lab_y = "", lab_title = "", y_reorder = TRUE) {
  # Purpose: wrap common NA histogram graphing for easy re-use
  # Input: in_df - dataframe, var - variable of interest, lod - level of detail (what to group by), 
  #       lab_x - x label, lab_y - y label, lab_title - chart title, y_reorder - Boolean, whether or not to sort
  #       x axis by descending y values
  # Output: histogram, formatted
  
  if(y_reorder) {
    in_df %>%
      group_by({{lod}}) %>%
      summarise(total_na = sum(is.na({{var}})), 
                total_obs = n(), 
                pct_na = total_na/total_obs) %>%
      ggplot(aes(x = reorder({{lod}}, -pct_na), y = pct_na)) +
      geom_bar(stat = 'identity') + 
      theme(axis.text.x = element_text(angle = 90)) +
      labs(x = lab_x, 
           y = lab_y,
           title = lab_title) +
      minimal_theme_GL
    } else {
    in_df %>%
      group_by({{lod}}) %>%
      summarise(total_na = sum(is.na({{var}})), 
                total_obs = n(), 
                pct_na = total_na/total_obs) %>%
      ggplot(aes(x = {{lod}}, y = pct_na)) +
      geom_bar(stat = 'identity') + 
      theme(axis.text.x = element_text(angle = 90)) +
      labs(x = lab_x, 
           y = lab_y,
           title = lab_title) +
      minimal_theme_GL
    }
}
```


Check Blank states
```{r check for blank states, }

# count_blank_na function was here

mobility_raw %>% 
  rename(state = sub_region_1, county = sub_region_2) %>%
  #select(state) %>% 
  count_blank_na(state)
```

Tidy up the data

1. rename some variables for ease of use
2. filter out observations where state = Blank (~400 values out of 1M)
  a. Note that there are no other columns to add context to help re-map
3. Only keep quantitative variables of interest - retail and recreation change, park change
4. Add a few new columns to separate out Year, month, and day of week (not for final csv, just to help in exploration)
5. Keep the columns of interest

*AR: why do we have missing states?*
*AR: why do we have missing retail_change?*
```{r rename some columns and filter out blanks}

# add_date_values function was here

mobility <- mobility_raw %>% 
  rename(state = sub_region_1, 
         county = sub_region_2, 
         retail_rec_change = retail_and_recreation_percent_change_from_baseline, 
         parks_change = parks_percent_change_from_baseline) %>%
  filter(state != "") %>% 
  add_date_values %>% # Only for initial exploration, not CSV output
  select(state, date, year, month, dow,
         # census_fips_code, county, place_id, # keeping for possible interim checks
         retail_rec_change, parks_change)
```

Check out NA's in quantitative variable
*AR: what do we do about states with such a high percentage of NA?*
```{r explore NAs by state and dates}
# Check # of NAs, and distribution
summary(mobility %>% select(retail_rec_change))
summary(mobility %>% select(parks_change))

# check # NAs 
for(i in c("retail_rec_change", "parks_change")) {
  print(i)
  mobility %>%
    count_blank_na(get(i))
}
```

Graph out some of these NAs, seeing if there are any patterns, or if they are random
```{r explore NAs by state and dates - graphical}
# plot_missing_histogram function was here

# Graphs for different types of NAs
state_NA <- mobility %>% 
  plot_missing_histogram(
                         # var = parks_change,
                         var = retail_rec_change, 
                         lod = state, 
                         lab_x = "State", 
                         lab_y = "% of values observations with NA", 
                         lab_title = "% of values missing by State")

month_NA <- mobility %>% 
  mutate(year_mo = paste(year, month, sep = "_")) %>%
  plot_missing_histogram(
                         # var = parks_change,
                         var = retail_rec_change, 
                         lod = year_mo, 
                         lab_x = "Year - Month", 
                         lab_y = "% of values observations with NA", 
                         lab_title = "% of values missing by Month",
                         y_reorder = FALSE)

first_NA <- mobility %>% 
  mutate(year_mo = paste(year, month, sep = "_")) %>%
  filter(year_mo %in% c("2020_02", "2020_03")) %>%
  plot_missing_histogram(
                         # var = parks_change,
                         var = retail_rec_change, 
                         lod = date, 
                         lab_x = "Date", 
                         lab_y = "% of values observations with NA", 
                         lab_title = "% of values missing by Date",
                         y_reorder = FALSE) +
  labs(subtitle = "Filtered to first two months with low % of NA")

dow_NA <- mobility %>% 
  plot_missing_histogram(
                         # var = parks_change,
                         var = retail_rec_change,
                         lod = dow, 
                         lab_x = "Day of Week", 
                         lab_y = "% of values observations with NA", 
                         lab_title = "% of values missing by Day of Week",
                         y_reorder = TRUE)

# Show graphs
state_NA
month_NA
first_NA
dow_NA
# (state_NA | month_NA) / 
#   (first_NA | dow_NA)
```
Old Charting Code (To delete)
```{r, echo = FALSE}
# state_NA <- mobility %>%
#   group_by(state) %>%
#   summarise(total_na = sum(is.na(retail_rec_change)), 
#             total_obs = n(), 
#             pct_na = total_na/total_obs) %>%
#   ggplot(aes(x = reorder(state, -pct_na), y = pct_na)) + 
#   geom_bar(stat = 'identity') + 
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "State", 
#        y = "% of values observations with NA",
#        title = "% of values missing by State") +
#   minimal_theme_GL

# month_NA <- mobility %>%
#   mutate(year_mo = paste(year, month, sep = "_")) %>%
#   group_by(year_mo) %>%
#   summarise(total_na = sum(is.na(retail_rec_change)), 
#             total_obs = n(), 
#             pct_na = total_na/total_obs) %>%
#   ggplot(aes(x = year_mo, y = pct_na)) +
#   geom_bar(stat = 'identity') +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "Year - Month", 
#        y = "% of values observations with NA",
#        title = "% of values missing by Month") +
#   minimal_theme_GL

# first_NA <- mobility %>%
#   mutate(year_mo = paste(year, month, sep = "_")) %>%
#   filter(year_mo %in% c("2020_02", "2020_03")) %>%
#   group_by(date) %>%
#   summarise(total_na = sum(is.na(retail_rec_change)), 
#             total_obs = n(), 
#             pct_na = total_na/total_obs, 
#             .groups = 'drop') %>%
#   ggplot(aes(x = date, y = pct_na)) +
#   geom_bar(stat = 'identity') +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "Date", 
#        y = "% of values observations with NA",
#        title = "% of values missing by Date", 
#        subtitle = "Filtered to first two months with low % of NA") +
#   minimal_theme_GL

# dow_NA <- mobility %>%
#   group_by(dow) %>%
#   summarise(total_na = sum(is.na(retail_rec_change)), 
#             total_obs = n(), 
#             pct_na = total_na/total_obs, 
#             .groups = 'drop') %>%
#   ggplot(aes(x = reorder(dow, -pct_na), y = pct_na)) +
#   geom_bar(stat = 'identity') +
#   theme(axis.text.x = element_text(angle = 90)) +
#   labs(x = "Day of Week", 
#        y = "% of values observations with NA",
#        title = "% of values missing by Day of Week") +
#   minimal_theme_GL
```

Comparing distribution of 0s across months to NAs across months
```{r comparing 0s to NAs}
mobility %>%
  filter(retail_rec_change == 0) %>%
  mutate(year_mo = paste(year, month, sep = "_")) %>%
  select(year_mo) %>%
  ggplot(aes(x = year_mo)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))


mobility %>%
  filter(is.na(retail_rec_change)) %>%
  mutate(year_mo = paste(year, month, sep = "_")) %>%
  select(year_mo) %>%
  ggplot(aes(x = year_mo)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))
```


Aggregating over Date and state, averaging all of the county `retail_rec_change`s
```{r aggregate by state two ways}
# Assuming NA's are bad, and we want to remove them
mobility_agg <- mobility %>%
  group_by(state, date) %>%
  # group_by(state, date, year, month) %>%
  summarise(avg_retail_rec_change = mean(retail_rec_change, na.rm = TRUE),
            avg_parks_change = mean(parks_change, na.rm = TRUE),
            .groups = 'drop') %>%
  select(state, date, avg_retail_rec_change, avg_parks_change)

# Assuming NA's are missing 0s, and we want to replace them with 0
mobility_agg_wNA <- mobility %>%
  replace_na(list(retail_rec_change = 0, parks_change = 0)) %>%
  group_by(state, date) %>%
  # group_by(state, date, year, month) %>%
  summarise(avg_retail_rec_change = mean(retail_rec_change),
            avg_parks_change = mean(parks_change),
            .groups = 'drop') %>%
  select(state, date, avg_retail_rec_change, avg_parks_change)
```
Check if all states have same number of observations
```{r check number of observations}
mobility_agg %>%
  group_by(state) %>%
  summarise(date_count = n(), 
            min_date = min(date), 
            max_date = max(date),
            date_range = max(date) - min(date),
            check = (max(date) - min(date)) - date_count = n())
```

Distribution of each quantitative variable
```{r check distributions}
mobility_agg %>%
  filter(state == "California") %>%
  
  add_date_values() %>%
  group_by(month) %>%
  summarise(measure = mean(avg_retail_rec_change)) %>%
  select(measure, month) %>%
  ggplot(aes(x = month, y = jitter(measure))) + 
  geom_point() + 
  labs(
    x = "change in retail/recreation",
    title = "Distribution of retail/recreation"
  ) +
  minimal_theme_GL

mobility_agg %>%
  filter(state == "California") %>%
  
  add_date_values() %>%
  group_by(month) %>%
  summarise(measure = mean(avg_parks_change)) %>%
  select(measure, month) %>%
  ggplot(aes(x = month, y = jitter(measure))) + 
  geom_point() + 
  labs(
    x = "change in retail/recreation",
    title = "Distribution of retail/recreation"
  ) +
  minimal_theme_GL


retail_hist <- mobility_agg %>%
  select(avg_retail_rec_change) %>%
  ggplot(aes(x = avg_retail_rec_change)) + 
  geom_histogram(bins = 50) + 
  labs(
    x = "change in retail/recreation",
    title = "Distribution of retail/recreation"
  ) +
  minimal_theme_GL

parks_hist <- mobility_agg %>%
  select(avg_parks_change) %>%
  ggplot(aes(x = avg_parks_change)) + 
  geom_histogram(bins = 50) + 
  labs(
    x = "change in parks",
    title = "Distribution of retail/recreation"
  ) +
  minimal_theme_GL

retail_hist / 
  parks_hist
```

```{r}
group_by({{lod}}) %>%
      summarise(total_na = sum(is.na({{var}})), 
                total_obs = n(), 
                pct_na = total_na/total_obs) %>%
      ggplot(aes(x = reorder({{lod}}, -pct_na), y = pct_na)) +
      geom_bar(stat = 'identity') + 
      theme(axis.text.x = element_text(angle = 90)) +
      labs(x = lab_x, 
           y = lab_y,
           title = lab_title) +
      minimal_theme_GL
```



Write transformed data for use in analysis
```{r}
write.csv(mobility_agg, 
# write.csv(mobility_agg_wNA,
           file = paste0("~/mids_w203_sp_21/lab_2-rbgs/data/interim/",
                          "mobility.csv"))
```

